


### Free memory Functions

```{r}
# Clear environment
rm(list = ls()) 

# Clear console
cat("\014")  # ctrl+L
```

#######################################################################################################
# Data exploration
#######################################################################################################
### Basic data exploration

#### Reading in and basic formatting of data

Start by reading in and formatting the data.

```{r}
autinsurance <- read.csv(file.choose("../Resources/autinsurance_V3.csv"), stringsAsFactors = TRUE)

```


# Libraries

We will use tidyverse libraries including ggplot2, tidyr, dplyr, and stringr to process this data.

We will also use gridExtra to be able to place ggplot2 plots side-by-side.

Also use caret and pROC when evaluating models.


```{r}
# packneeded <- c('ggplot2','stringr','tidyr','dplyr', 'gridExtra', 'caret', 'pROC', "psych" , "moments")
# install.packages(packneeded, dependencies = TRUE)
```


```{r load-libs, echo=TRUE, eval=TRUE,message=FALSE,warning=FALSE}
library(ggplot2)
library(stringr)
library(tidyr)
library(dplyr)
library(gridExtra)
library(caret)
library(pROC)
library(psych)
library(moments)
```



```{r}
names(autinsurance) # variable names
```


#######################################################################################################
## Assessing Normality 
#######################################################################################################

Histograms

Normality can be visually assessed with a histogram. A quick histogram can be made using base R’s hist() function:

```{r}
hist(autinsurance$months_as_customer)
#ggplot(data=autinsurance) + geom_histogram(aes(x=months_as_customer) , fill = "red", bins = 30)
```


##  Density (Curve) Plots
You can also use a density plot to visually assess normality:

```{r}
#plot(density(autinsurance$months_as_customer, na.rm=T))
ggplot(data=autinsurance) + geom_density(aes(x=months_as_customer) , fill = "red")
```

## QQ Plots
A QQ plot can also be used to assess normality:

A QQ plot is a plot of the actual data values against a normal distribution (which has a mean of 0 and standard deviation of 1).
```{r}
# qqnorm(autinsurance$months_as_customer)
ggplot(autinsurance, aes(sample = months_as_customer)) + 
  geom_qq() + geom_qq_line()
```

##  Skewness and Kurtosis
There are a number of ways to get values for skewness and kurtosis. You can get the values alongside other stats with the psych packmonths_as_customer’s describe(). You can also use skewness() and kurtosis() functions from the moments packmonths_as_customer:

```{r}
#describe(autinsurance)
```

```{r}
# install.packmonths_as_customers("moments")
library(moments)
skewness(autinsurance$months_as_customer)
kurtosis(autinsurance$months_as_customer)
```

##  Tests of Normality

Univariate tests of normality are also easily done in R. The stats packmonths_as_customer (included with R) contains both the Kolmogorov-Smirnov (K-S) test and the Shapiro-Wilk’s test. The Shapiro-Wilk’s test is more common and easier to compute in R:

```{r}
shapiro.test(autinsurance$months_as_customer)
```

The above p-value is significant at less than 0.05, signifying non-normality.


#######################################################################################################
## Transforming Variables
#######################################################################################################

The following as some of the many methods you can use if you wish to deal with non-normality by transforming variables (DataNovia, n.d.):

## Square Roots

-- for positive skew
```{r}
skewness(autinsurance$months_as_customer)  # Raw data
skewness(sqrt(autinsurance$months_as_customer))
```

-- for negative skew
```{r}
skewness(autinsurance$months_as_customer)
skewness(sqrt(max(autinsurance$months_as_customer+1)-autinsurance$months_as_customer))
```

## Log

-- for positive skew
```{r}
skewness(autinsurance$months_as_customer)
skewness(log10(autinsurance$months_as_customer))
```

-- for negative skew
```{r}
skewness(autinsurance$months_as_customer)
skewness(log10(max(autinsurance$months_as_customer+1)-autinsurance$months_as_customer) )
```


## Inverse

-- for positive skew 
```{r}
skewness(autinsurance$months_as_customer)
skewness(1/autinsurance$months_as_customer)
```

-- for negative 
```{r}
skewness(autinsurance$months_as_customer)
skewness(1/max((autinsurance$months_as_customer)+1))
```


#######################################################################################################
# Summary of Transform Methods
#######################################################################################################
Below is a quick summary of all of the transform methods supported in the method argument of the preProcess() function in caret.

“BoxCox“: apply a Box–Cox transform, values must be non-zero and positive.
“YeoJohnson“: apply a Yeo-Johnson transform, like a BoxCox, but values can be negative.
“expoTrans“: apply a power transform like BoxCox and YeoJohnson.
“zv“: remove attributes with a zero variance (all the same value).
“nzv“: remove attributes with a near zero variance (close to the same value).
“center“: subtract mean from values.
“scale“: divide values by standard deviation.
“range“: normalize values.
“pca“: transform data to the principal components.
“ica“: transform data to the independent components.
“spatialSign“: project data onto a unit circle.
The following sections will demonstrate some of the more popular methods.


# Scale
The scale transform calculates the standard deviation for an attribute and divides each value by that standard deviation.

```{r}
# load libraries
library(caret)
# calculate the pre-process parameters from the dataset
preprocessParams <- preProcess(autinsurance, method=c("scale"))
# summarize transform parameters
print(preprocessParams)
# transform the dataset using the parameters
transformed <- predict(preprocessParams, autinsurance)
# summarize the transformed dataset
summary(transformed)
```

# Center
The center transform calculates the mean for an attribute and subtracts it from each value

```{r}
# load libraries
library(caret)
# calculate the pre-process parameters from the dataset
preprocessParams <- preProcess(autinsurance, method=c("center"))
# summarize transform parameters
print(preprocessParams)
# transform the dataset using the parameters
transformed <- predict(preprocessParams, autinsurance)
# summarize the transformed dataset
summary(transformed)
```

# Standardize
Combining the scale and center transforms will standardize your data. Attributes will have a mean value of 0 and a standard deviation of 1.

```{r}
# load libraries
library(caret)
# calculate the pre-process parameters from the dataset
preprocessParams <- preProcess(autinsurance, method=c("center", "scale"))
# summarize transform parameters
print(preprocessParams)
# transform the dataset using the parameters
transformed <- predict(preprocessParams, autinsurance)
# summarize the transformed dataset
summary(transformed)
```

# Normalize
Data values can be scaled into the range of [0, 1] which is called normalization.

```{r}
# load libraries
library(caret)
# calculate the pre-process parameters from the dataset
preprocessParams <- preProcess(autinsurance, method=c("range"))
# summarize transform parameters
print(preprocessParams)
# transform the dataset using the parameters
transformed <- predict(preprocessParams, autinsurance)
# summarize the transformed dataset
summary(transformed)
```

# Box-Cox Transform
When an attribute has a Gaussian-like distribution but is shifted, this is called a skew. The distribution of an attribute can be shifted to reduce the skew and make it more Gaussian. The BoxCox transform can perform this operation (assumes all values are positive).

```{r}
# load libraries
library(mlbench)
library(caret)

# calculate the pre-process parameters from the dataset
preprocessParams <- preProcess(autinsurance, method=c("BoxCox"))
# summarize transform parameters
print(preprocessParams)
# transform the dataset using the parameters
transformed <- predict(preprocessParams, autinsurance)
# summarize the transformed dataset (note pedigree and age)
summary(transformed)

```
Notice, we applied the transform to only two attributes that appear to have a skew. Running the recipe, you will see:

# Yeo-Johnson Transform
Another power-transform like the Box-Cox transform, but it supports raw values that are equal to zero and negative.

```{r}
# load libraries
library(mlbench)
library(caret)
# load the dataset
data(PimaIndiansDiabetes)
# summarize pedigree and age
summary(autinsurance)
# calculate the pre-process parameters from the dataset
preprocessParams <- preProcess(autinsurance, method=c("YeoJohnson"))
# summarize transform parameters
print(preprocessParams)
# transform the dataset using the parameters
transformed <- predict(preprocessParams, autinsurance)
# summarize the transformed dataset (note pedigree and age)
summary(transformed)
```



